---
title: "ESM 204 Section Exercise Week 9"
author: "Dustin Duncan"
format: 
  html:
    code-fold: true
    toc: true
    number-sections: true
    embed-resources: true
editor: visual
execute:
  echo: true
  message: false
  warning: false
---

```{r}
#| message: false
#| warning: false
rm(list = ls())
library(tidyverse)
library(pracma)
library(stats)
```

## Risk Preferences:

### Risk neutral:

$$
\mathrm{E[U(x)] = U(E[x])} \\
\mathrm{洧녷_1洧녣(洧논_1) + 洧녷_2洧녣(洧논_2) = 洧녣(洧녷_1洧논_1 + 洧녷_2洧논_2)}
$$ 

**Indifferent between the risky payout versus the expected value with certainty**

### Risk Averse: If U(x) is concave or

$$
\mathrm{E[U(x)] < U(E[x])} \\
\mathrm{洧녷_1洧녣(洧논_1) + 洧녷_2洧녣 (洧논_2) < 洧녣(洧녷_1洧논_1 + 洧녷_2洧논_2)}
$$ 

**Prefer having the expected value with certainty versus taking the gamble**

### Risk Loving: If U(x) is concave or

$$
\mathrm{E[U(x)] > U(E[x])} \\
\mathrm{洧녷_1洧녣(洧논_1) + 洧녷_2洧녣 (洧논_2) > 洧녣(洧녷_1洧논_1 + 洧녷_2洧논_2)}
$$

**Prefer taking the risky payout versus the expected value with certainty**

## Exercise 1

Consider a gamble:

-   pays x_1 = 0 with p = 0.75

-   pays x_2 = 2 with p = 0.25

-   Utility given by:

$$
\mathrm{U(x)=2x-0.5x^{2}}
$$

-   

    (a) Graph the utility function with wealth on the horizontal axis and utility on the vertical axis

Where x is wealth:

```{r}

utility = function(x){
  U = 2*x - 0.5*(x^2)
  return(tibble(x=x, U=U))
}

x = seq(from=0,to=150,by=1)

test = utility(x)

ggplot(test, aes(x, U)) + 
  geom_point()
```

-   

    (b) Is this utility function linear, convex or concave?

-   

    (c) Add the utility of having x1 = 0 and x2 = 2

-   

    (d) What is the expected value of the gamble E\[x\]?

-   

    (e) What is the utility of having the expected value of the gamble with certainty U(E(x))

-   

    (f) What is the expected utility of the gamble, E(U(x))?

-   

    (g) plot (e) and (f) on your graph

-   

    (h) Is an individual with this utility function risk neutral, risk averse or risk seeking? Demonstrate this on your graph. How could you tell from part (a) alone?
